---
title: "Genome assemblies query"
output:
  html_notebook:
    toc: yes
    toc_depth: 3
    # toc_float:
    #   collapsed: false
    #   smooth_scroll: false
---

## Setup
```{r}
library(RSQLite)
library(dbplyr)
library(dplyr)
library(DBI)
```

```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), dbname = "C:\\Users\\Kelly\\Documents\\Python\\Organisms\\Organismdb.db")

```

Check if connection worked
```{r}
attributes(con)
dbListTables(con)
dbExistsTable(con,"Organisms")

dbListFields(con,"Organisms")
```

View table!
```{r}
tbl(con,"Organisms")
```

Get row count:
```{r}
tally(tbl(con,"Organisms"))

#Show this query as SQL request:
show_query(tally(tbl(con,"Organisms")))
```


Function to query assemblies
```{r}
query<- function(nx) {
x<-dbSendQuery(con, 
               'SELECT * 
             FROM Organisms WHERE ScientificName == ?')
dbBind(x,list(nx))
z<-(dbFetch(x,n=1)) #n=1 specifies results to return, doesn't work as well if multiple results are fetched
dbClearResult(x) #VERY important! VERY!
return(z)
dbClearResult(x) #VERY important! VERY!
}
```

Query assemblies and append new columns from SQLite database
```{r}
db_results<-data.frame()  #make a new data.frame

for (i in assemblies[,"Organism"]) {
  x<-query(i)
  if (nrow(x)==0) {
    x[nrow(x)+1,]<- c(rep("NULL",11))}
  db_results<-rbind(db_results,x)}

#Bind new columns to assemblies dataframe
assemblies_complete<-cbind(assemblies,db_results)
```

If there are any entries missing in db_results, db_results won't cbind to assemblies(differing numbers of rows)
```{r}
which(assemblies[,"Organism"] %starts% "Drosophila melanogaster")
```


Check which rows are missing:
```{r}
'%!in%' <- function(x,y)!('%in%'(x,y))
which(assemblies[,"Organism"] %!in% db_results[,"ScientificName"])

which(db_results[,"ScientificName"] %!in% assemblies[,"Organism"])
```

Verify that Organism matches Scientific Name in all rows
```{r}
which(assemblies_complete[,"Organism"] != assemblies_complete[,"ScientificName"])
```

Rows 100,200 and 400 weren't in the database!
Print out the missing rows so I can go back to the Python script and add the missing data to the db.
```{r}
assemblies_complete[c(100,200,400),]
```

All done!




